---
title: "string_practice"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

## **Task 3: Practice with strings**

-   Install the `datapasta` package (if you don't have it already)

-   Restart R

-   Create a new .Rmd in your `R` folder called `string_practice.Rmd`

-   Copy all contents of the html table below to your clipboard:

```{r}
alarm_report<- data.frame(
  stringsAsFactors = FALSE,
              date = c("2020-03-14", "2020-03-15", "2020-04-10", "2020-04-18"),
          building = c("Engineering-North",
                       "Bren Hall","Engineering-South","Engr-North"),
     alarm_message = c("10:02am – HVAC system down, facilities management alerted",
                       "8:24am – Elevator North out of service",
                       "12:41am – Fire alarm, UCSB fire responded and cleared",
                       "9:58pm – Campus point emergency siren, UCPD responded")
)
```

-   Back in your `string_practice.Rmd`, create a new code chunk

-   With your cursor in your code chunk, go up to Addins in the top bar of RStudio. From the drop-down menu, choose 'Paste as data frame'. Make sure to add code to store the data frame as `alarm_report`

-   Practice working with strings by writing code to update `alarm_report` as follows (these can be separate, or all as part of a piped sequence):

    -   Replace the "Engr" with "Engineering" in the `building` column

    ```{r}
    alarm_report %>% 
      mutate(name_new = str_replace(building, pattern = "Engr", replacement = "Engineering"))
    ```

    -   Separate the `building` column into two separate columns, `building` and `wing`, separated at the dash

    ```{r}
    alarm_report %>% 
      separate(building,into = c("building","wing"), sep = "-")
    ```

    -   Only keep observations with the word "responded" in the `alarm_message` column

    ```{r}
    alarm_report %>% 
      filter(str_detect(alarm_message, pattern = "responded"))
    ```

    -   Separate the message *time* from the rest of the message by separating at `--`

    ```{r}
    alarm_report %>% 
      separate(alarm_message,into = c("time","alarm_message"), sep = "–")
    ```

    -   Convert the date column to a Date class using `lubridate`

    ```{r}
    alarm_report %>% 
      mutate(date = lubridate::ymd(date))
    ```
